---
title: "Exploratory figures"
author: "Andrew MacDonald"
date: "April 8, 2016"
output: pdf_document
---

```{r echo=FALSE, message=FALSE}
## note that this works only interactively -- ie it won't work if you just click "Knit"
## load all data from main project
remake::dump_environment()
```

## block-by-block analysis of the data

The idea here is that the relationship we predict could be found at the level of individual blocks, as well. This would allow us to conduct an analysis of effect sizes. 

```{r}

## a list of blocks, with a "taxa" and "factors" element of each
f <- . %>% 
  lapply(split, f = .[["factors"]][["Block"]]) %>% 
  transpose()

invert_zoops <-  
  list(inverts = inverts_tts_fin %>% f,
       zoops   = zoops_tts_fin %>% f,
       bact    = bact_tts_fin %>% f)


## presence-absence data
presabs <- function(df) {
  for (i in names(df)) {
    df[[i]] <- as.numeric(df[[i]] > 0)
  }
  df
}

invert_zoops_pa <- invert_zoops %>% 
  ## remove all zero species within each block:
  at_depth(2, . %>% map_at("taxa", ~.x[,(colSums(.x > 0) > 0)])) %>% 
  ## converte inverts and zoops to presence absence
  map_at(c("inverts", "zoops"), . %>% 
           map(. %>% 
                 map_at("taxa", presabs)))

```

So far so good. Now let's build the PERMANOVA response:

```{r message=FALSE}
AdonisData <- function(Data, ...){  
  adonis(Data[["taxa"]] ~ species , data = Data[["factors"]], ...)
}

rdaData <- function(Data, ...){  
  rda(Data[["taxa"]] ~ species , data = Data[["factors"]], ...)
}

adonis_list <- invert_zoops_pa %>% 
  at_depth(2, AdonisData, method = "euclid")

rda_list <- invert_zoops_pa %>% 
  at_depth(2, rdaData)


block_level_rsqs <- adonis_list %>% 
  map(. %>% map_dbl(~.x[["aov.tab"]][["R2"]][[1]])) %>% 
  map_df(~data_frame(Block = names(.x), rsq = .x), .id = "grp")

kable(block_level_rsqs)

# ## and f-vals
# block_level_fs <- adonis_list %>% 
#   map(. %>% map_dbl(~.x[["aov.tab"]][["F.Model"]][[1]])) %>% 
#   map_df(~data_frame(Block = names(.x), fval = .x), .id = "grp")
```



### plot the results

```{r}
block_lvl_rs_ordered <- block_level_rsqs %>%
  mutate(grp = ordered(grp, levels = c("bact", "zoops", "inverts")))

block_lvl_rs_ordered %>% 
  ggplot(aes(x = grp, y = rsq, group = Block)) +
  geom_point() + 
  geom_line()
```

Perhaps not surprisingly, when we break down the data to very small samples, we find very noisy results with no clear pattern. 

## number of species in common

first let's see what species are present everywhere

```{r}
ubiquity <- invert_zoops %>% 
  at_depth(2, . %>% map_at("taxa", ~.x[,(colSums(.x > 0) > 0)])) %>% 
  at_depth(2, . %>% map_at("taxa", ~ data_frame(ubiq = sum(colSums(.x > 0) == 6),
                                                nspp = ncol(.x)))) %>% 
  map_df(. %>% map_df("taxa", .id = "Block"), .id = "taxa")
```

OK let's try to plot that

```{r}
ubiquity %>% 
  ggplot(aes(x = nspp, y = ubiq, colour = taxa)) +
  geom_point() + 
  stat_smooth(method = "lm") +
  facet_wrap(~taxa, scales = "free")
```


## correlation analysis

could we analyze figure 3 directly, with a linear model? the goal here is to determine how much of the effect is truly robust.

```{r}

test_slope_robustness <- function(inverts_adonis_ini, inverts_adonis_fin,
                              zoops_adonis_ini, zoops_adonis_fin,
                              bact_adonis_ini, bact_adonis_fin, REPS){
  
  sizes <- data_frame(taxa = c("bact", "zoops", "inverts"),
                      size = 1:3)
  # could replace these values with actual size later
  
  observed <- r2_plot_df_maker(inverts_adonis_ini, inverts_adonis_fin,
                               zoops_adonis_ini, zoops_adonis_fin,
                               bact_adonis_ini, bact_adonis_fin) %>% 
    left_join(sizes) %>% 
    group_by(time) %>% 
    do(broom::tidy(lm(number ~ size, data = .)))
  
  ### randomization tests
  
  null_sims <- replicate(n = REPS, {
    ss <- 1:30 %>% 
      split(f = gl(5, 6)) %>% 
      lapply(sample) %>% 
      unlist
    
    inverts_tts_ini$factors$species <- inverts_tts_ini$factors$species[ss]
    zoops_tts_ini$factors$species <- zoops_tts_ini$factors$species[ss]
    bact_tts_ini$factors$species <- bact_tts_ini$factors$species[ss] 
    
    inverts_tts_fin$factors$species <- inverts_tts_fin$factors$species[ss]
    zoops_tts_fin$factors$species <- zoops_tts_fin$factors$species[ss]
    bact_tts_fin$factors$species <- bact_tts_fin$factors$species[ss] 
    
    inverts_adonis_ini <- AdonisData(inverts_tts_ini, method = I("euclid"))
    zoops_adonis_ini <- AdonisData(zoops_tts_ini, method = I("euclid"))
    bact_adonis_ini <- AdonisData(bact_tts_ini, method = I("euclid"))
    inverts_adonis_fin <- AdonisData(inverts_tts_fin, method = I("euclid"))
    zoops_adonis_fin <- AdonisData(zoops_tts_fin, method = I("euclid"))
    bact_adonis_fin <- AdonisData(bact_tts_fin, method = I("euclid"))
    
    r2_plot_df_maker(inverts_adonis_ini, inverts_adonis_fin,
                     zoops_adonis_ini, zoops_adonis_fin,
                     bact_adonis_ini, bact_adonis_fin) %>% 
      left_join(sizes) %>%
      group_by(time) %>% 
      do(broom::tidy(lm(number ~ size, data = .)))
    
    
  }, simplify = FALSE)
  
  obs_slope <- observed %>% filter(term == "size") %>% select(time, estimate)
  
  null_sims_summary <- null_sims %>% 
    bind_rows(.id = "sim") %>% 
    filter(term == "size") %>% 
    select(time, estimate) %>%
    bind_rows(obs_slope) %>% 
    left_join(obs_slope %>% 
                rename(obs = estimate)) %>% 
    group_by(time, obs) %>% 
    summarize(null_mean = mean(estimate),
              null_sd   = sd(estimate),
              null_p    = sum(estimate >= obs) / length(estimate)) %>% 
    mutate(ses = (obs - null_mean) / null_sd)
  
  return(list(null_sims, null_sims_summary))
}

null_sims_spear %>% 
  bind_rows(.id = "sim") %>% 
  filter(term == "size") %>% 
  .[["estimate"]] %>% 
  hist(xlim = c(-0.05, 0.15), col = "lightgrey", main = "null distribution of slopes") %>% 
  abline(v = obs_slope, col = "darkgreen",  lwd = 2)
```
