---
title: "The effect of habitat variation on organisms of different body size"
author: "A. Andrew M. MacDonald, Diane S. Srivastava, Vinicius F. Farjalla (et al?)"
output:
  md_document: default
  html_document: default
  pdf_document: default
  word_document:
    reference_docx: reference.docx
bibliography: /home/andrew/Documents/Mendeley_reference_lists/@MS_habitat_size_org.bib
---

#Introduction
Previous observational studies [@Farjalla2012] have shown that larger
bodied organisms show a greater degree of habitat filtering: there is greater
environmental structure as body size increases from bacteria (weak),
zooplankton (intermediate) and insects (strong). This is especially
interesting as these different environments are caused by variation in habitat
preference among several bromeliad species -- some species are found under
shrub cover, while others are meters away in full sun.

Sun and shade habitats have different species compositions. This may be caused
by organism' s local environmental tolerances; sensitivity to heat may vary
among the three groups of organisms studied.  However, species compositions may
also be affected by the effects of temperature fluctuations on metabolism.
Warmer animals may show increased feeding rate, which may increase top-down
regulation and alter both density and richness of lower trophic levels.

In this experiment we will test these observations by homogenizing these
communities, removing variation among communities as much as possible.  We
hypothesize that as habitat filtering acts, via mortality and -- for non-
metamorphosing organisms -- reduced fecundity, we will observe a 'relaxation'
of the community towards the orginal species composition.  We hypothesize that
this relaxation will be greatest for insects, less for zooplankton and nearly
zero for bacteria.

## Methods

We performed two complementary experiments to test how different groups within
these communities respond to environmental variation.  In each experiment, we
removed the communities from bromeliads, sampled them, and then returned a
homogenized community to all plants.  Plants were replaced in a similar local
habitat to their original, and enclosed in a mesh cage to prevent
colonization. After a 12 day period, we sampled communities again and tested
for a 'relaxation' in composition towards their original composition.  In our
first experiment (among-species) we tested how sensitivy to different
bromeliad species.  Our second experiment (within-species) examined the
effects of habitat independently of species, by comparing communities within
the same bromeliad species in open- and closed-canopy habitats.

We performed these experiments at the Parque Nacional de Jurubatiba, in Rio de
Janeiro state, Brazil ($22^{\circ}$ S $41^{\circ}$).  For our between--species
experiment we used _Aechmea nudicaulis_, _Neoregelia cruenta_, and _Vriesea
neoglutinosa_; our between --habitat experiment used _Neoregelia cruenta_
only, sampled from exposed and shaded habitats.  _N. cruenta_ is both common
and highly plastic, occuring in a broader range of habitats than any other
bromeliad in this study site.

Macroinvertebrate communities were sampled by washing the plants throughly and
seiving the water through a (dimenions) sieve.  They were identified to the
lowest taxonomic level possible in the field (genus or family).  Homogenized
communities were the arithemetic mean of the densities of each group in
original communities (though often quite a bit lower because of natural
mortality).

To homogenize zooplankton, protist and bacteria communities, we collected and
equal volume of inoculum from each plant, mixed and returned equal volumes to
all bromeliads.  In our first experiment we used 100ml of inoculum from each
plant, because _A. nudicaulis_ is very small and this was the maximum volume
we could obtain while simultaneously selecting average-sized plants.



## Results
## List of figures

```{r message=FALSE,warning=FALSE,echo=FALSE,results='hide'}
# load packages -----------------------------------------------------------

library(reshape)
library(reshape2)
library(grid)
library(ggplot2)
library(vegan)
library(igraph)
library(plyr)
library(lubridate)
library(knitr)

source("~/Dropbox/PhD/ibutton.functions/ibutton.functions.R")
source("../R.scripts/habitatExp.R")
opts_chunk$set(dev="png",results='hide',message=FALSE,warning=FALSE,echo=FALSE)
```

```{r fig.cap=""} 
#newer_plot_select_blocks() 
threespp <- plot_select_blocks_arrow()
## starting points
p1 <- ggplot(data=threespp,aes(x=CA1.initial,y=CA2.initial,colour=brom_sp))+geom_point(,size=5)
## homogenized points (different colour)
p1 <- p1 + geom_point(aes(x=CA1.homogenized,y=CA2.homogenized),size=5)
p1 <- p1+geom_segment(aes(x=CA1.homogenized,xend=CA1.final,y=CA2.homogenized,yend=CA2.final,colour=brom_sp),arrow=arrow(length=unit(0.2,"cm")))

#svg("insects.svg",width=11.2,height=8.4)
p1+theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))+facet_wrap(~block,scales="free")
#dev.off()
``` 


```{r}
threesppzoo <- plot_select_blocks_arrow(Dataframe=zoop,homogen_insects=FALSE)
p2 <- ggplot(data=threesppzoo,aes(x=CA1.initial,y=CA2.initial,colour=brom_sp))+geom_point()
## homogenized points (different colour)
p2 <- p2 + geom_point(aes(x=CA1.final,y=CA2.final))
p2 + facet_grid(~brom_sp)

## final and initial
## melt data
threesppzoo_melt <- melt(threesppzoo,id.vars=c("block","brom_sp","bromeliad","habitat"))
newcols <- sapply(as.character(threesppzoo_melt$variable),strsplit,split=".",fixed=TRUE)
threespp_newcol <- data.frame(threesppzoo_melt,do.call(rbind,newcols))

threesppzoo_new <- threespp_newcol[!names(threespp_newcol)%in%c("variable")]
threesppzoo_new <- rename(threesppzoo_new,c("X1"="axis","X2"="sampling"))
new_zoo <- dcast(threesppzoo_new,formula=block+brom_sp+bromeliad+habitat+sampling~axis,value.var="value")

svg("zoop.svg",width=11.2,height=8.4)
ggplot(new_zoo,aes(x=CA1,y=CA2,shape=sampling,colour=brom_sp))+geom_point(size=5)+facet_wrap(~block,scales="free")+theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))
dev.off()


head(zoop)

dlply(zoop,.(bromeliad),dcast,sampling~spp)

zoo.nspp <- count(zoop,.(bromeliad,sampling))

## that can be graphed right away:
bromeliad2 <- bromeliad[,c(1:2,6)]
bromeliad2 <- rename(bromeliad2,c("Brom"="bromeliad"))
zoo_n_spp <- join(zoo.nspp,bromeliad2,by="bromeliad")
test <- dcast(data=zoo_n_spp,species+Block+bromeliad~sampling,value.var="freq")

ggplot(test,aes(x=final,y=initial,colour=species))+geom_line()#+facet_grid(~Block)

```

### Figure 1  Insect community
change over time in the inter-species experiment.  The graph is based on a non-
metric multidimensional scaling of a euclidian distance matrix among
communities.  Each dot represents the initial (observed) community composition
of each bromeliad; dots are coloured according to species.  We formed
'homogenized' communities by allocating six bromeliads (two of each species) to
blocks and dividing the combined invertebrate community evenly among them.
Arrows show the change from this identical starting community to the composition at the end of the experiment (position of the arrow head).A test of our
hypothesis requires these arrows (representing community 'relaxation' back to
the original composition) be both in a nonrandom direction (ie tending to point
to intial compositions) *and* longer for insects than for zooplankton, which in
turn show larger change than bacteria.





```{r fig.cap=""}
ggplot(sarahjane,aes(x=Date.Time,y=Value,colour=habitat),group=ibutton)+geom_line()
```
### Figure 2
The temporal variation in temperature between two habitats -- open and closed.  The effect of a very cold day can be seen on the right.

```{r eval=FALSE}
## doesn't work right :(  needs a complete check.
twohab <- plot_select_blocks_arrow(block_vector=c("SarahJane","Martha","Rose","Donna","Pond"))
ggplot(data=twohab$first,aes(x=CA1,y=CA2,colour=habitat))+geom_point()+geom_point(data=twohab$homo)+geom_segment(data=twohab$arrows,aes(x=CA1.homo,xend=CA1.final,y=CA2.homo,yend=CA2.final,colour=habitat.final),arrow=arrow(length=unit(0.2,"cm")))
```


## Literature Cited

