---
title: "The effect of habitat variation on organisms of different body size"
author: "A. Andrew M. MacDonald, Vinicius F. Farjalla , Diane S. Srivastava"
fontsize: 12pt
output:
  pdf_document:
    includes:
      in_header: formatting/myheader.tex
    latex_engine: xelatex
  md_document: default
  html_document: default
  word_document:
    reference_docx: formatting/reference.docx
geometry: margin=1in
bibliography: orgsize.bib
csl: formatting/the-american-naturalist.csl
---

```{r echo=FALSE}
fig_nums <- captioner()
tab_nums <- captioner(prefix = "Table")
fig <- pryr::partial(fig_nums, display = "cite")
tab <- pryr::partial(tab_nums, display = "cite")
```


# Abstract

At the local scale, species composition of often varies among habitat patches in
response to differences in environmental variables. These differences act as
"filters": organisms cannot occur in a habitat patch when their sensitivity to
local environmental conditions prevents their establishment or persistence.
However, not all groups of organisms are equally sensitive to the same
environmental variables. Observational studies have already suggested that
larger-bodied species are more strongly filtered by habitat than smaller- bodied
species. However, this mechanism has never been tested experimentally. In our
study we placed highly similar aquatic communities (composed of insects,
zooplankton and bacteria) into different environments and observed subsequent
divergence in composition. We predicted that insects would show a strong change
in composition, zooplankton a moderate change and bacteria little to no change.
Our results are consistent with these predictions, with the exception of a
weaker signal in insects than expected. These results suggest that organism size
could provide a framework for integrating studies of environmental filtering
across different scales and systems.

# Introduction

models of community assembly explain how local species composition are assembled from the regional species pool. They can be divided into neutral and niche-based processes. Neutral processes are based on the demographic equivalence of all species. the relative abundance of species in any local patch is dependent on their frequency in the regional species pool, paired with stochastic demographic processes such as immigration and emigration. In contrast, niche-based models of community assembly emphasize the trait differences among species. In these models, species are able to coexiest if they posses sufficeniently different traits (stabilizing processes) and their relative abundance is due to differences in their fitness (e.g. survival and reproduction). Niche-based models emphasize how the presence of species in a site is correlated with the abiotic and biotic conditions of that site. These determine whether a species is able to colonize or persist in a given patch. In niche-based, as in neutral models, variance in community composition is also derived from stochasticity in demographic processes. Both categorite of model also make predictions about how patches differ from each other in species composition (beta diversity). In a neutral model, differences between patches are caused by stochasticity in demographic processes, particularly dispersa.  In a niche-based model this variation is increased by variation among habitat patches, which selects for different species compositions in each. Separating the different patterns produced by these two methods usually involves variance partitioning. This approach has its detractors. But the question remains for ecologists: when and under what circumstances will local communities be assembled by neutral or niche-based processes? 

 In one of the first studies across different groups of
organisms in the same community, Farjalla et al. [-@Farjalla2012] demonstrated that body
size is a key trait determining environmental signal in community composition. They conducted an observational study across many different organism types, spanning orders of magnitude in body size, along a single environmental gradient. They considered two different hypotheses regarding how increasing body size should change the balance of niche-based vs neutral community assembly processes. The size-dispersal and size-plasticity hypotheses. The former hypothesis suggests that larger animals have weaker dispersal -- therefore more stochasticity in community composition -- while smaller organisms have such high dispersal rates that variation in composition must be driven by the suitability of the local environment. In contrast the size-plasticity hypothesis suggests that more important than the ability to disperse is the differences in the local environment. That is, smaller organisms should be less sensitive to environmental variation (ie more plastic) and larger organisms more so. They found that groups of organisms which are physically larger show a greater degree of
habitat filtering: there is weaker environmental structure as body size
decreases from insects (strong signal), zooplankton (moderate signal) and bacteria
(weak signal). Their analysis of observational data found support for the size-plasticity hypothesis.

Why should the strenght of environmental filtering decrease with organism size? there are two possible mechanisms that may make communities of smaller organisms more strongly filtered. At the individual level, organisms which are very small maybe able to find microhabitats in a way that mitigates varation abmong patches. smaller organisms also have simpler metabolisms. Bacteria are often able to consume a wide variety of carbon sources. Bacteria can even use environmental DNA, or transfer plasmids among unrelated cells, acquiring new functions. 

The individual-level effects suggested by the size-plasticity hypothesis are difficult to test because of population level processes: especially dispersal. At the population level, smaller organisms may also show less of a relationship with environmental variation. if smaller organisms disperse more rapidly, then we would expect to see populations of smaller species maintained, even in patches where they have negative population growth rates (i.e. mass effects). smaller organisms have higher growth rates, higher rates of turnover, and shorter generation times. This suggests that they could persiste in highly variable or fluctuation habitats, purely by demographic processes. Also, they could stick aroun by evolution, as they have larger effects population sizes (does mutation rate show an effect with regard to population size?). rapid evolution to antiobotics is well know.
In both cases, therefore, space is an issue -- the processes of dispersal can either exaggerate or diminish the effects seen by Farjalla et al. 

However, the work of Farjalla et al suggests a manipulative experiment: to
remove variation among patches, while preserving abiotic variation. If these
environmental differences. We studied the same groups of organisms (insects,
zooplankton and bacteria) in the same patchy, highly variable habitat (three
species of tank bromeliads). We created communities that were as similar as
possible in composition, then exposed these communities to their original
environmental conditions.



We test three different predictions of the size-plasticity hypothesis: 

Question 1: communities which begin at a simiar species composition will diverge, but this divergence
will be greatest between larger species and increasingly weak in smaller
organisms

Question 2: This increase in beta diversity will correlate with environmental differences; this correlation will be stronger for larger organisms.

Question 3: The specific taxa responding to environmental variation will possess traits that are related to their ability to sustain highly variable environments.




# Methods

## Experimental design

We used te bromeliad
species from different habitats ("between-species"), while the second
contrasted the same bromeliad species from different habitats ("within-
species"). We focused on three bromeliad species from three
contrasting habitats; these species and habitats are identical to those surveyed
by Farjalla et al.  [-@Farjalla2012]. Each bromeliad species was collected from its typical
habitat: _Aechmea nudicaulis_ (full sun habitats), _Vriesea neoglutinosa_
(partial shade), and _Neoregelia cruenta_ (full shade). _Neoregelia_ has a uniquely large habitat range at this site, occuring in both full shade and full sun; only shade plants were used in this study.

### Study site

We performed our experiments at the Parque Nacional de Jurubatiba, Northeast Rio
de Janeiro state, Brazil ($22^{\circ}$ S $41^{\circ}$). The landscape is
composed of large semi-open patches of shrubs which form shady patches
interspersed with very exposed, sunny habitats (_restinga_ vegetation).
Bromeliads (Bromeliaceae) form dense colonies within this landscape. The
landscape is composed of large semi-open clumps of shrubs interspersed with very
exposed, open sand. Several bromeliad taxa here collect standing water
(phytotelmata) in deep leaf axils, and are known to contain diverse communities 
of macroinvertebrates, zooplankton and bacteria.

## Experimental design

Both between- and within-species experiments used similar designs: five blocks
of six bromeliads each. To create each block in the between-species experiment,
we collected two bromeliads of each of the three species; in the within-species experiment we
collected three bromeliads from open habitats and three from closed habitats.  In both
experiments, all bromeliads in a block were stocked with a nearly identical
starting community, intermediate between that of all six bromeliads in that
block. We made these starting communities by sampling, then combining, each of
three major groups: macroinvertebrates, zooplankton and bacteria. 

We sampled macroinvertebrates by thoroughly rinsing each bromeliad and passing
the water through 1mm and 180μm mesh to collect aquatic macroinvertebrates.  We
identified these animals to morphospecies, then divided them equally among the
six bromeliads in the block. 

We sampled the starting zooplankton and bacteria communities by mixing an equal
volume of filtered tank water from each plant (approximately 100ml plant^-1^),
returning the same volumes to all bromeliads. 

We emptied bromeliads by hanging them upside down to dry for at least 24 hours
and then rinsing each plant with 70% ethanol. Any coarse detritus found in the
bromeliads was similarly cleaned, frozen and thawed (to kill any macroinvertebrates eggs). We
confirmed that this technique removed all invertebrates and most detritus by
dissecting an empty bromeliad. Bromeliads were placed in a local habitat similar
to their original, each enclosed in a 2mm mesh cage to prevent colonization by
insects. We then added the starting communities of macroinvertebrates,
zooplankton and bacteria.

The experiment lasted for 12 days, after which we collected all
macroinvertebrates, and a sample of zooplankton and bacteria.
Macroinvertebrates were identified while live and preserved in ethanol.
Zooplankton were collected on 50-micron Nytex mesh and fixed in 5% buffered
formalin. Subsamples were then identified to the lowest
taxonomic unit possible (species in most cases, except for bdelloid rotifers
and harpaticoid copepods, left at the level of class and order, respectively). 
We measured bacterial community composition using denaturing gradient gel
electrophoresis (DGGE; Muyzer et al. 1993). This analysis furnishes an
approximation of bacterial diversity in the form of Operational Taxonomic
Units (OTUs). We based these collection methods as closely as possible upon
previous observational work, to facilitate a direct comparison  [@Farjalla2012].

### Analyses


For each of our two experiments, we performed a pair of statistical tests: a
permutation test (PERMANOVA) to detect differences in the degree of
environmental signal among groups, and a multivariate regression of abundance
to identify those taxa which contributed most to the differences. PERMANOVA
measures the amount of difference in community composition between treatment
groups and compares this to the expected distribution under a null hypothesis of no treatment effects.
If larger-bodied groups of organisms are more filtered, then we expect to find
a greater difference in community composition emerge among the
macroinvertebrates, less among the zooplankton, and little to none among
bacteria.

## Tables


`r tab_nums("disp_tab", "this is the table of disperions")`
```{r results='asis'}
disp_difs_long  %>% 
  lm(value ~ taxa * timing, data = .) %>% 
  stargazer(header = FALSE)

```

### another table
```{r results='asis'}
disp_aov <- lmer(value ~ taxa * timing + (1|block), data = disp_difs_long)

stargazer::stargazer(disp_aov, header = FALSE)
```

```{r echo=FALSE}
getF <- function(x) sprintf("%.2f", x[["aov.tab"]][["F.Model"]][[1]])
getP <- function(x) x[["aov.tab"]][["Pr(>F)"]][[1]]
getR <- function(x) sprintf("%.2f", x[["aov.tab"]][["R2"]][[1]])
```

`r tab_nums("perm_3sp", "Results from PERMAOVAs run on each of the three groups of organisms, both before and after the experiment. Both F-ratios and R^2^ values are higher for inverts than they are for zooplankton or bacteria")`

|                    |        | F                            | p                            | R^2^                         |
| -------- | ------------------| ---------|--------- | ---------- |
| macroinvertebrates | before | `r getF(inverts_adonis_ini)` | `r getP(inverts_adonis_ini)` | `r getR(inverts_adonis_ini)` |
|                    | after  | `r getF(inverts_adonis_fin)` | `r getP(inverts_adonis_fin)` | `r getR(inverts_adonis_fin)` |
| zooplankton        | before | `r getF(zoops_adonis_ini)` | `r getP(zoops_adonis_ini)` | `r getR(zoops_adonis_ini)` |
|                    | after  | `r getF(zoops_adonis_fin)` | `r getP(zoops_adonis_fin)` | `r getR(zoops_adonis_fin)` |
| bacteria           | before | `r getF(bact_adonis_ini)` | `r getP(bact_adonis_ini)` | `r getR(bact_adonis_ini)` |
|                    | after  | `r getF(bact_adonis_fin)` | `r getP(bact_adonis_fin)` | `r getR(bact_adonis_fin)`                             |


## Figures

![disp](../figures/plot_disp_taxa.pdf)
`r fig_nums("disp", "behold the figure of dispersion")`

![between](../figures/Fig1.pdf)
`r fig_nums("bw_spp", "The x-axis has the group studied. the Y axis the r2 of the data from a PERMANOVA. It turns out that a organisms get smaller, the effect of time on the species composition of the data gets weaker")`

![within](../figures/Fig2.pdf)
`r fig_nums("ftwo", "this would seem to be figure 2")`

## Results  

Goal 1: dispersion differences among the different organism groups.  

We found that dispersion, in all cases, was lower at the end of the experiment than at the beginning. This indicates that our treatments were effectingve in removing beta diverstiy among patches. that is, that all habitats really did end up being very similar at the beginning, as we would expect. However, the amount of this decline was the same across all organism groups (`r tab("disp_tab")`), as indicated by the 



Our interspecific experiment found a large effect of bromeliad species on the community composition of invertebrates, then less and less for zooplankon and bacteria (`r fig_nums("bw_spp")` )


We found more signal in insects and zooplankton relative to bacteria. These differences can be seen in 

Can we get an effect size out of the insects relative to the zooplankton and bacteria? we could, instead of using PERMANOVAs with block as a random factor, we could actually just test this effect size metric on animals from large to small with ordinal contrasts. 

Another possibility: is there a convenient way to measure the differences between the initial and final community compositions. 




## Discussion

We did discover different amounts of environmental signal ithe different groups -- more in invertebrates and zooplankton than in bacteria. In the direction that we expected, based on the size-sensitivity hypothesis. We found support for these same patterns when we examined individual species responses -- more species of insects showed a change in response to environmental variation than did zooplankton. No bacteria OTU ever showed a predictable difference between before and after. Just as Farjalla et al predicted, the strength of environmental fitlering is in negative proportion to body size. Over the same time interval, insects and zooplankton developed more of a response to the environment than did bacteria. 

The difference between insects and zooplankton was not as great as we expected it would be. that was probably because we were missing the active vs passive differences. Insect females choose where to place their eggs, evalutating resources and predation risk (citecitecite). On the other hand, the macroinvertebrate taxa that did respond are also the more abundant taxa; it may be that we only had the ability to detect patterns for these taxa. Both of these issues at once (should we do permanova on mortality?). Both size does predict some of the variation in response to the environment. Since active dispersers are also larger, there is then a relationship between that life history and body size -- larger animals are more likely to disperse actively. It may be that those groups show a response in that direction as well. 

In conclusion, here we have shown that a hypothesis made by variance partitioning and by observations, can be supported by a manipulative experiment and a different statistical approach. that, I guess, is probably good news for variance partitioning methods, which have come under a lot of criticism lately. iN our experiment we removed the variation in community composition between habitats, and were essentially able to watch the recovery of beta diversity in our system. Measured as both turnover and dispersion, in that test I haven't even done yet. 

In other words our experiment has not only confirmed some aspects of the body size habitat senstivity experiment, but also investigated the rate at which beta diversity is recovered for these different groups. p 

## References

Farjalla, V. F., Srivastava, D. S., Marino, N. a C., Azevedo, F. D., Dib, V.,
Lopes, P. M., Esteves, F. A. (2012). Ecological determinism increases with
organism size. Ecology, 93(7), 1752–9.

Muyzer G., de Waal E.C., Uitterlinden A.G. (1993). Profiling of complex
microbial populations by denaturing gradient gel electrophoresis analysis of
polymerase chain reaction-amplified genes coding for 16S rRNA. Applied
Environmental Microbiology, 59(3), 695-700.

