---
title: "Supplementary Material"
author: "A. Andrew M. MacDonald; Diane S. Srivastava; Vinicius"
output:
  pdf_document:
    latex_engine: xelatex
    includes:
      in_header: formatting/myheader_suppmatt.tex
    toc: true
  md_document: default
  html_document: default
  word_document:
    reference_docx: reference.docx
geometry: margin=1in
fontsize: 12pt
---


```{r echo=FALSE}
Sfig_nums <- captioner()
Stab_nums <- captioner(prefix = "Table")
Sfig <- pryr::partial(Sfig_nums, display = "cite")
Stab <- pryr::partial(Stab_nums, display = "cite")
```

# Exploratory figures

![](../figures/inverts.pdf)
![](../figures/zoop.pdf)
![](../figures/bact.pdf)

# Dispersion

mixed effect ANOVA table
```{r results='asis'}
anova(disp_mixed) %>% 
    knitr::kable()
```

residual plot for the dispersion model
```{r}
plot(disp_mixed)
```

![](../figures/one_one_plot.pdf)
`r Sfig_nums("one2one", "This is the different mean differences to centroid on the same plot. shown is the 1:1 line.")`

# PERMANOVAS

```{r, echo = FALSE}
get_adonis_table <- function(permanova_answer){
  permanova_answer[["aov.tab"]] %>% as.data.frame() %>% knitr::kable()
}
```


## initial

### inverts
```{r results = 'asis'}
get_adonis_table(inverts_adonis_ini)
```

### zooplankton
```{r results = 'asis'}
get_adonis_table(zoops_adonis_ini)
```

### bacteria
```{r results = 'asis'}
get_adonis_table(bact_adonis_ini)
```

## final
### insects
```{r results = 'asis'}
get_adonis_table(inverts_adonis_fin)
```

### zooplankton
```{r results = 'asis'}
get_adonis_table(zoops_adonis_fin)
```

### bacteria
```{r results = 'asis'}
get_adonis_table(bact_adonis_fin)
```

# multivariate GLM

### results of multivariate GLMS

```{r}
plot(inverts_manyglm_fin)
```


```{r}
plot(zoops_manyglm_fin)
```

```{r}

plot(bact_manyglm_fin)
```

## manyglm


test for interactions
```{r}
#anova(inverts_manyglm_fin, inverts_manyglm_fin_add)

```

